---
title: "Countries Codebook"
author: "Duu Renn" # Change this
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document: 
    number_sections: yes
    toc: yes
---

```{r include=FALSE, cache=FALSE}
# We'll talk about these options later. 
require(knitr)
opts_chunk$set(
  tidy=TRUE,     # display code as typed # Changed to true because my comments are too long
  tidy.opts=list(width.cutoff=75),
  size="small",    # slightly smaller font for code
  echo=TRUE,
  results='markup',
  strip.white=TRUE,
  fig.path='figs/fig',
  cache=FALSE,
  highlight=TRUE,
  width.cutoff=132, 
  size='footnotesize',
  out.width='.9\\textwidth',
  fig.retina=FALSE,
  message=FALSE,
  comment=NA)
```

# The Countries Dataset
The 'Countries' dataset is a smaller version of a dataset by Pippa Norris. The [codebook](http://www.jakebowers.org/Data/Countries/Democracy%2520Crossnational%2520Codebook%2520March%25202009.pdf) has variable descriptions, but we will briefly practice describing these variables some more here so that you are able to use Rmarkdown with Political Science data.  

# Load the Dataset
```{r}
load(url("http://www.jakebowers.org/Data/countries.rda"))
```

# Code: What variables are contained in this dataset? 
```{r}

```

# Explain: Pick a variable and past its codebook description here. 

# Code: Describe the contents of the variable appropriately. 
```{r}

```

# Code: Provide a figure that helps describe the distribution of the variable. 
```{r}

```


# Complete and Review
- Compile this document using the 'knit' option in RStudio. 
- Submit a pull request. 


