---
title: 'Exploration 0: Statistical Inference Review'
author: "Jake Bowers"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document:
    toc: no
  html_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    number_sections: yes
    toc: yes
geometry: margin=1in
graphics: yes
mainfont: Minion Pro
fontsize: 10pt
---

```{r opts, include=FALSE, cache=FALSE}
require(knitr)

opts_chunk$set(
  tidy=TRUE,
  tidy.opts=list(width.cutoff=75),
  size="small",
  echo=TRUE,
  results='markup',
  strip.white=TRUE,
  fig.path='figs/fig',
  cache=FALSE,
  highlight=TRUE,
  width.cutoff=132, 
  size='footnotesize',
  out.width='.9\\textwidth',
  fig.retina=FALSE,
  message=FALSE,
  warning=FALSE,
  comment=NA)
```

```{r initialize,echo=FALSE}
##First, just setup the R environment for today:
if(!file.exists('figs')) dir.create('figs')

options(SweaveHooks=list(fig=function(){
			   par(mar=c(3.5, 3, 1.1, 0),
			       pty="s",
			       mgp=c(1.5,0.5,0),
			       oma=c(0,0,0,0))},
			 echo=function(){options(continue=" ") ##Don't show "+" prompts,
			 options(prompt=" ")
			 }),
	digits=4,
	scipen=8,
	width=132
	)
options(error=function(){options(prompt="> ",continue="+ ");NULL})
```

Imagine that an old friend calls. She is now in charge of Countering Violent
Extremism and Improving Civil Society programs at the United Nations. She
tells you that she happened to hear that you are a specialist in political
science from a mutual friend and that she needs your help. She has been asked
to write a report called "Ten Years after the London Bombing of 2005", and,
although she has hired some data analysts, she is having trouble understanding
exactly what they are doing. You gather that, in fact, at least one of her
data analysts has left the UN and your high powered UN official friends has an
important report due but has no technical help.

Here is what she asks:

Did the London Subway and Bus bombings of July 2005 change the civic
activity of people in the UK? The UK Home Office had a survey in the field
in 2005. Because of confidentiality requirements she can only share a sample
of 1000 of these people.

She sends you some code that she says she found after, "they confiscated her
machine". You ask, "Who confiscated whose machine?" And your friend replies,
"I can't say. But can you help? I need to understand what is going on here."


```{r}
load(url("http://jakebowers.org/Data/ho05.rda"))
```

\begin{small}
\begin{verbatim}
  ### CODEBOOK
  postbomb: 1=interviewed after the bombing, 0=interviewed before the bombing

  grphrs: 6.1.1 Which of the following groups, clubs or organisations
    have you been involved with during the last 12 months? That's anything
    you've taken part in, supported, or that you've helped in any way, either
    on your own or with others. Please exclude giving money and anything that
    was a requirement of your job.

    6.1.2 In the last 12 months have you given unpaid help to any groups, clubs or
    organisations in any of the following ways?

    6.1.5 Approximately how many hours have you spent helping this/these group(s),
    club(s) or organisation(s) in the past 4 weeks?

  infhrs: In the last 12 months have you done any of the following things,
    unpaid, for someone who was not a relative?

    This is any unpaid help you, as an individual, may have given to other people,
    that is apart from any help given through a group, club or organisation. This
    could be help for a friend, neighbour or someone else but not a relative.

    6.4.4 Now just thinking about the past 4 weeks. Approximately how many hours
    have you spent doing this kind of thing/these kind of things in the past 4
    weeks?

  hlphrs: grphrs+infhrs
\end{verbatim}
\end{small}


```{r}
## Aarrgh make data with no missing data on outcomes
wrkdat<-ho05[!is.na(ho05$hlphrs),]
```

After you ponder this codebook, you ask, "Why did your analyst put 'aaargh'
into the code?" Your friend says, "That is not important. When I asked my data
analyst to tell me the effect of the bombings on civic life in the UK, she
gave me this next piece of code and output. How does this answer my question? I didn't ask for a
coefficient!! Can you show me how to get this number in a way that is simpler?
Why use averages anyway? What is going on here?"

```{r}
effectlm<-lm(hlphrs~postbomb,data=wrkdat) ## the linear model
coef(effectlm) ## coefficients
```

She continues, "One of my subordinate always brags about his statistics
abilities at parties. So I asked him to help me interpret
this number, he said, `Well, you can't trust this coefficient because your
dependent variable is not Normal.' I didn't understand and so, the next day I
asked my main analyst to explain. She didn't get a chance to explain but the
agents recovered this from her computer:"


```{r, eval = F}
## Outcome distribution
summary(wrkdat$hlphrs)
plot(density(wrkdat$hlphrs))
rug(wrkdat$hlphrs)
```

"Why would the analyst or those managing her as an asset want to know about
the distribution of voluntarism?"

You ask, "What do you mean...`those managing her as an asset'?"

She says, "Oh! I...I... meant that she was a real asset to me as a manager in
my organization. Anyway, what about the question whether people in the UK are
Normal or not normal or whatever? Why should I care about this? Why shouldn't
I want the original number that I got in the first place (now that I
understand what was going on)? By the way, when I asked this question of two
different analysts, I got two different hunks of code. I get the sense that
there are two issues with the `normal/not-normal' question. Also, I am being
driven crazy by people sending me emails with R code in them. Help!"

"Here is the code from Analyst 1. What is the issue that he is responding to?
Why more coefficients here? How are these coefficients telling us about
effects?"

```{r}
# install.packages("quantreg")
require(quantreg)
effectRq<-rq(hlphrs~postbomb,data=wrkdat,tau=.5) # tau=.5 is middle quantile
coef(effectRq) ## coefficients
```

"Here is the code from Analyst 2: What is she doing? Why? What do her comments
about a `CLT' mean?"

```{r}
permuteFn<-function(y){
  shuffledBombing<-sample(wrkdat$postbomb)
  meandiff<-coef(lm(y~shuffledBombing))[["shuffledBombing"]]
  return(meandiff)
}
```

```{r, cache=T}
set.seed(20150101)
results<-replicate(10000,permuteFn(y=wrkdat$hlphrs))
mean(results)

```

```{r, fig.width=8, fig.height=4, out.width=".5\\textwidth", message=T, warning=T, eval = F}
par(mfrow=c(1,2),oma=rep(0,4),mgp=c(1.5,.5,0),pty="s")
plot(density(results/sd(results)),xlab="Mean Diffs")
curve(dnorm(x,sd=1,mean=0),from=-4,to=4,col="blue",add=TRUE) ## CLT ok.
obscoef<-coef(effectlm)[[2]]
rug(obscoef,lwd=3,ticksize=.5)

qqnorm(results/sd(results))
qqline(results/sd(results))
```

```{r, eval = F}
pnorm(obscoef,sd=1,mean=0,lower.tail=FALSE) # p-value
2*(min( c( pnorm(obscoef,sd=1,mean=0),1-pnorm(obscoef,sd=1,mean=0) )))
summary(effectlm)$coef[2,c(1,4)]

```

"Thanks so much for all of your help! Now I do see that the mention of
`Normal' referred to two different problems. Now I have one last question:  "I
also asked my analyst to tell me how people in London acted compared to people
in other parts of the UK. After all, the bombings occurred in London. But I
got more crazy coefficients. Can you help? Can you show me how to get these
numbers without using any commands like `lm`?"

```{r}
wrkdat$london<-factor(wrkdat$GOR=="London",labels=c("Not London","London"))
## test the recode
with(wrkdat,table(london,GOR=="London",useNA="ifany"))

effectsBySubGroupsLm<-lm(hlphrs~london*postbomb,data=wrkdat)
coef(effectsBySubGroupsLm)
```

